%!TEX root = ../thesis.tex

\thispagestyle{myheadings}

\chapter{\g for Spin-1/2 Particles}
\label{gDirac}

Do I even want to include this derivation? Needs work if I am to do so. (If I remove it remove the reference in chapter 1 and just reference schwartz directly.)

Here is provided a relatively straightforward calculation of the g-factor for spin-1/2 particles, simplified from a derivation by Schwartz \cite{Schwartz}. Start with the Dirac equation in the presence of an external electromagnetic field: 
	\begin{align}
		(i\gamma^{\mu}\partial_{\mu} - m -e \gamma^{\mu}A_{\mu})\psi = 0,
	\label{eq:dirac}
	\end{align}
where $\psi$ is the wavefunction of a spin-1/2 particle, $m$ is its mass, $A_{\mu}$ is the electromagnetic field operator, $\partial_{\mu}$ is the 4-gradient, and $\gamma^{\mu}$ are the gamma matrices\footnote{Here $\hbar$ and c will be set equal to 1.}. Multiplying \equref{eq:dirac} by itself but with the sign on $m$ switched, one arrives at the equation
	\begin{align}
		[(i\partial_{\mu}-eA_{\mu})(i\partial_{\nu}-eA_{\nu})\gamma^{\mu}\gamma^{\nu} - m^{2}]\psi = 0.
	\end{align}
This can be split this into its symmetric and antisymmetric parts,
	\begin{align}
		&\Big(\frac{1}{4}\{i\partial_{\mu}-eA_{\mu},i\partial_{\nu}-eA_{\nu}\}\{\gamma^{\mu}, \gamma^{\nu}\} \notag \\
		&+\frac{1}{4}[i\partial_{\mu}-eA_{\mu},i\partial_{\nu}-eA_{\nu}][\gamma^{\mu}, \gamma^{\nu}] - m^{2}\Big)\psi = 0,
	\label{eq:symmantisymm}
	\end{align}
where $[,]$ and $\{,\}$ stand for the commutator and anti-commutator relations respectively. Using the identities 
	\begin{align}
		\frac{i}{2}[\gamma^{\mu}, \gamma^{\nu}] = \sigma^{\mu\nu}
	\end{align}
and
	\begin{align}
		[i\partial_{\mu}-eA_{\mu},i\partial_{\nu}-eA_{\nu}] = -ie[\partial_{\mu}A_{\nu} - \partial_{\nu}A_{\nu}] = -ie F_{\mu\nu},
	\end{align}
where $\sigma^{\mu\nu}$ is a tensor with elements consisting of the Pauli matrices, and $F_{\mu\nu}$ is the electromagnetic field tensor, \equref{eq:symmantisymm} can be formed as
	\begin{align}
		((i\partial_{\mu}-eA_{\mu})^{2} - \frac{e}{2} F_{\mu\nu}\sigma^{\mu\nu} - m^{2})\psi = 0.
	\end{align}
Expanding out the tensor objects
	\begin{align}
		\frac{e}{2} F_{\mu\nu}\sigma^{\mu\nu} = -e
		\begin{pmatrix}
			(\vec{B} + i \vec{E})\cdot \vec{\sigma} && 0 \\
			0 && (\vec{B} - i \vec{E})\cdot \vec{\sigma} \\
		\end{pmatrix}
\end{align}
and forming a new covariant derivative
	\begin{align}
		\slashed{D}^{2} = D_{\mu}^{2} + \frac{e}{2} F_{\mu\nu}\sigma^{\mu\nu}
	\end{align}
where $D_{\mu} = \gamma^{\mu}\partial_{\mu}$ is your ordinary covariant derivative, by moving to momentum space (this part is awkward) you can arrive at the equation
	\begin{align}
		\frac{(H+eA)^2}{2m}\psi = (\frac{m}{2} + \frac{(\vec{p}+e\vec{A})^{2}}{2m} -2 \frac{e}{2m}\vec{B}\cdot\vec{s} \pm i \frac{e}{m}\vec{E}\cdot\vec{s})\psi.
	\end{align}
Lo and behold, you have arrived at the Dirac g = 2 result, contained in front of the magnetic piece in the form of Equation \ref{eq:magneticmoment}.

How then does such a term change at loop level? Most generally the vertex of a particle interacting with a magnetic field through the mediation of a photon can be represented by 
	\begin{align}
		iM^{\mu} = \bar{u}(q_{2}) (f_{1}\gamma^{\mu} + f_{2}p^{\mu} + f_{3}q^{\mu}_{1} + f_{4}q^{\mu}_{2}) u(q_{1})
	\end{align}
where $q_{1}$ and $q_{2}$ are the ingoing and outgoing four-momenta respectively, which can be constrained on-shell, and p is the four-momenta of the photon, which is off-shell. The $f_{i}$ are in general contractions of four-momenta and gamma matrices. By using the Gordon identity
	\begin{align}
		\bar{u}(q_{2}) (q^{\mu}_{1} + q^{\mu}_{2}) u(q_{1}) \notag \\
		= (2m)\bar{u}(q_{2}) \gamma^{\mu}u(q_{1}) + i\bar{u}(q_{2})\sigma^{\mu\nu}(q^{\nu}_{1} - q^{\nu}_{2}) u(q_{1})
	\end{align}
any Feynman diagram can be reorganized into the form 
	\begin{align}
		iM^{\mu} = (-ie)\bar{u}[F_{1}(\frac{p^{2}}{m^2})\gamma^{\mu} + \frac{i\sigma^{\mu\nu}}{2m} p_{\nu}F_{2}(\frac{p^{2}}{m^2})]u,
	\end{align}
where $F_{1}$ and $F_{2}$ are form factors. One notices that the $F_{2}$ piece is reminiscent of our magnetic dipole moment form that we derived from the Dirac equation. So the problem now becomes for any Feynman diagram calculation in any theory, at any order, to solve for this $F_{2}$ to determine the contribution to the magnetic dipole moment.
